---
title: "Polynomial Simulation like in LRD paper"
author: "Adam C Sales & Ben B Hansen & Josh Wasserman"
date: "August 2022"
output:
  github_document:
    html_preview: false
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=FALSE,message=FALSE,warning=FALSE)
knitr::opts_knit$set(root.dir = '.')
```


```{r getCommandArgs,include=FALSE}
aaa=commandArgs(TRUE)
aaa=gsub("--","",aaa)
for(a in aaa) eval(parse(text=a))
```

General dependencies.
```{r}
library('knitr')
library('kableExtra')
library('testthat')
source('displaySim.r')
devtools::load_all("../..")
```


## Running Simulation

Initialization. In order to re-run the simulations, set the `nreps`
variable to a positive integer corresponding to the desired number of replications before initiating this script.
Note that `nreps=0` corresponds to no simulations,
just print results from previously saved simulations.



```{r setNreps,echo=TRUE, warning=FALSE, message=FALSE}
if (!exists('nreps') ){
  nreps <- 0
} else if(!is.numeric(nreps)) error("non-numeric nreps detected in workspace")

if(nreps){
  message(paste("Running simulation with",nreps,"replications"))
}else message("Making table from previously-run simulation")

```

### Parallel Computing

In order to run the simulation in parallel, if `nclust` is set to a positive integer, set the `nclust` variable to a positive integer corresponding to the desired number of clusters.
Note that `nclust=1` corresponds to running the simulation sequentially.

```{r parallel, eval=nreps>0}
clust=FALSE
if(exists("nclust")) if(is.numeric(nclust)) if(nclust>1) if(require(parallel)) {
  cl <- makeCluster(nclust)
  clust=TRUE
}
```

### Running the simulation

```{r runSimulation, eval=nreps>0}
source('./runSimulations.R')
```

## Outputting the Results in a Table


The following gives the results in Table 4 of the paper, in addition
to the break-down of RMSE into bias and variance, and analogous
results for normally-distributed errors.

```{r make table}
load("totalPolySim.Rdata")
tab <- prntTab(totalPoly,5,full=TRUE)
#rownames(tab) <- rep(c('level','RMSE','bias','sd'),6)
colnames(tab) <- gsub('(flex|ols)\\.se.','deg=',colnames(tab))#c(rep(paste0('deg=',1:4),2),'')
# colnames(tab)[ncol(tab)] <- 'n/a' # this is for LLR column
kable(tab,format='html',caption='Full results for polynomial simulation',digits=2) %>%
  kable_styling() %>%
  column_spec(6,border_right=TRUE) %>%
  column_spec(11,border_right=TRUE) %>%
  # add_header_above(c(" " = 1, "flexida" = 5, "OLS" = 5, "Loc. Lin." = 1))%>%
  add_header_above(c(" " = 1, "flexida" = 5, "OLS" = 5))%>%
  #group_rows("$t_3$ Error",1,12)%>%group_rows("N(0,1) Error",13,24)%>%
  group_rows("linear",1,5) %>%
  group_rows('antiSym',6,10) %>%
  group_rows('sine',11,15) #%>% group_rows("linear",13,16)%>%group_rows('antiSym',17,20)%>%group_rows('oneSide',21,24)
```

```{r coverage bound}
N_PAPER_SIMS <- 5000
MIN_PAPER_COVERAGE <- 0.93
N_TEST_SIMS <- ncol(totalPoly[[1]])
MIN_COVERAGE_BOUND <- MIN_PAPER_COVERAGE - 2 * sqrt(0.95 * 0.05 / N_PAPER_SIMS) -
  2 * sqrt(0.95 * 0.05 / N_TEST_SIMS)
MIN_COVERAGE_BOUND
```

```{r run tests}
testthat::expect_true(all(rownames(tab)[c(5, 10, 15)] == "ciCover"))
testthat::expect_true(
  all(tab[5, 1:5] > MIN_COVERAGE_BOUND)
)
testthat::expect_true(
  all(tab[10, 3:5] > MIN_COVERAGE_BOUND)
)
testthat::expect_true(
  all(tab[15, 3:5] > MIN_COVERAGE_BOUND)
)
```

Session information
```{r session}
sessionInfo()
```
