---
title: "propertee vs. geex"
author: "Joshua Wasserman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("geex")
)
```

```{r install, message = FALSE, warning = FALSE}
library(propertee)
library("geex")
```

## Overview
We provide basic validation for the model-based sandwich standard errors provided in \texttt{propertee} by comparing them to estimates from the \texttt{geex} package (Saul and Hudgens, 2020). In two examples, one with and one without clustering, we estimate the standard error of an intent-to-treat (ITT) effect estimate that has been adjusted for covariates using a prior covariance model. In the M-estimation framework, the prior covariance model defines one set of estimating equations, and the covariate-adjusted intent-to-treat effect estimate defines another. One can stack them togther and use the \texttt{geex} package to obtain joint estimates of the model coefficients and ITT effect, or one can fit the covariance model first, then use the \texttt{propertee} package to feed necessary information about the first-stage estimating equations to variance estimation for the ITT effect estimate. We demonstrate the equivalence of these two appraoches below.

## Example with non-clustered data
```{r nonclustered}
set.seed(980)
n <- 60
cmod_df <- data.frame(
  "uid" = seq_len(n),
  "z" = c(rep(0, 3 * n/4), rep(1, n/4)),
  "x1" = rnorm(n, sd = 2),
  "x2" = rep(c(rep(0, n/4), rep(1, n/4)), 2)
)
eps <- rnorm(nrow(cmod_df))
theta <- c(75, 0.1, -0.5, 2.5)
cmod_df$y <- model.matrix(~ x1 + x2 + z, cmod_df) %*% theta + eps

cmod_df$in_q <- c(rep(0, n/2), rep(1, n/2))
q_df <- cmod_df[cmod_df$in_q == 1,]

cmod <- lm(y ~ x1 + x2, cmod_df)
des <- rct_design(z ~ unitid(uid), q_df)
damod <- lmitt(y ~ 1, design = des, data = q_df, weights = ate(des), offset = cov_adj(cmod))

estFun <- function(data){
  function(theta) {
    # covariance model eqns
    Xstar <- model.matrix(y ~ x1 + x2, data)
    cmod_eqns <- drop(data$y - Xstar %*% theta[1:3]) * Xstar
    
    # itt model eqns
    X <- model.matrix(y ~ x1 + x2, data)
    Z <- model.matrix(y ~ z, data)
    if (data$in_q == 1) {
      damod_eqns <- drop(data$weight * (data$y - X %*% theta[1:3] - Z %*% theta[4:5])) * Z
    } else {
      damod_eqns <- rep(0, 2)
    }
    
    out <- c(cmod_eqns, damod_eqns)
    return(out)
  }
}

geexRes <- geex::m_estimate(estFun,
                            data = cbind(cmod_df,
                                         "weight" = c(rep(1, n/2), ate(des, data = q_df))),
                            root_control = geex::setup_root_control(start = rep(0.1,5)))
print(paste("geex estimate of var(tau_hat):", round(geexRes@vcov[5,5], 8)))
print(paste("propertee estimate of var(tau_hat):",
            round(vcovDA(damod, type = "CR0", cadjust = FALSE)[2,2], 8)))
```

## Example with clustered data
```{r clustered}
set.seed(50)
nc <- 4
mi <- 20
cmod_df <- data.frame(
  "cid" = rep(seq_len(4), each = mi),
  "uid" = rep(seq_len(mi), nc),
  "z" = c(rep(0, 3 * mi), rep(1, mi)),
  "x1" = rbinom(nc * mi, 1, 0.5),
  "x2" = rnorm(nc * mi)
)

theta <- c(75, 7.5, -1, 2.5)
error_sd <- round(runif(nc, 1, 3), 1)
icc <- 0.2
eps <- rnorm(nrow(cmod_df))
Sigma <- matrix(0, nrow = nrow(cmod_df), ncol = nrow(cmod_df))
for (i in (seq_len(nc) - 1)) {
  msk <- cmod_df$cid == (i + 1)
  Sigma[msk, msk] <- diag(error_sd[i + 1] - icc, nrow = sum(msk)) + icc
}
A <- chol(Sigma)
eps <- t(A) %*% eps
cmod_df$y <- model.matrix(~ x1 + x2 + z, cmod_df) %*% theta + eps

cmod_df$in_q <- c(rep(0, mi), rep(1, 3 * mi))
q_df <- cmod_df[cmod_df$in_q == 1,]

cmod <- lm(y ~ x1 + x2, cmod_df)
des <- rct_design(z ~ cluster(cid), q_df)
damod <- lmitt(y ~ 1, design = des, data = q_df, weights = ate(des), offset = cov_adj(cmod))

clusterEstFunc <- function(data){
  function(theta) {
    # covariance model eqns
    Xstar <- model.matrix(y ~ x1 + x2, data)
    cmod_agg_func <- ifelse(dim(Xstar)[2] > 1, colSums, sum)
    cmod_eqns <- cmod_agg_func(drop(data$y - Xstar %*% theta[1:3]) * Xstar)
    
    # itt model eqns
    X <- model.matrix(y ~ x1 + x2, data)
    Z <- model.matrix(y ~ z, data)
    damod_agg_func <- ifelse(dim(Z)[2] > 1, colSums, sum)
    if (unique(data$in_q) == 1) {
      damod_eqns <- damod_agg_func(
        drop(data$weight * (data$y - X %*% theta[1:3] - Z %*% theta[4:5])) * Z)
    } else {
      damod_eqns <- rep(0, 2)
    }
    
    out <- c(cmod_eqns, damod_eqns)
    return(out)
  }
}

geexRes <- geex::m_estimate(clusterEstFunc,
                            data = cbind(cmod_df,
                                         "weight" = c(rep(1, mi), ate(des, data = q_df))),
                            units = "cid",
                            root_control = geex::setup_root_control(start = rep(0.1,5)))

print(paste("geex estimate of var(tau_hat):", round(geexRes@vcov[5,5], 8)))
print(paste("propertee estimate of var(tau_hat):",
            round(vcovDA(damod, type = "CR0", cadjust = FALSE)[2,2], 8)))
```

## References
Saul BC, Hudgens MG (2020). “The Calculus of M-Estimation in R with geex.” Journal of Statistical Software, 92(2), 1–15. doi: 10.18637/jss.v092.i02.