---
title: "Introduction to propertee"
author: "propertee Authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to propertee}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
devtools::load_all("~/repositories/_r/propertee/")
#library(propertee)
```

## Main Features

The **propertee** package, Prognostic Regression Offsets with Propagation of
ERrors (for Treatment Effect Estimation), facilitates direct adjustment for
experiments and observational studies with design-informed standard errors and
flexible options for covariance adjustment. It uses explicit specification of
study design to provide probability of assignment weights and standard errors
that appropriately reflect the design. For covariance adjustment of its Hajek
and (one-way) fixed effects estimates, it enables offsetting the outcome against
predictions from a dedicated covariance model, with standard error calculations
propagating error as appropriate from the covariance model.

The main workflow consists of two main steps and one optional step:

1. Generate a `Design` object which encodes the study design, including the unit
   of assignment, treatment status of each unit of assignment, and optionally
   block information. This is accomplished with the `obs_design()`,
   `rct_design()` or `rd_design()` functions.
2. Optionally, fit a covariate adjustment model.
3. Fit a model to estimate a treatment effect, accounting for the design
   information, and optionally the covariate adjustment. This is done via the
   `lmitt()` function.

## Example Data

The example dataset comes from the state of Tennessee's Student-Teacher
Achievement Ratio (STAR) experiment. Students were randomly assigned to three
possible classroom conditions: small (13 to 17 students per teacher), regular
class (22 to 25 students per teacher), and regular-with-aide class (22 to 25
students with a full-time teacher's aide).

```{r}
data(STARdata)
table(STARdata$stark)
```

For simplicity for this first example, we will examine a single binary
treatment - "small" classrooms versus "regular" and "regular+aide" classrooms.

```{r}
STARdata$starkbinary <- STARdata$stark == "small"
table(STARdata$starkbinary)
```

After this basic example, we will see how **propertee** makes it easy to handle
non-binary treatment variables by introducing `dichotomy`s.

The outcome of interest is a reading score at the end of kindergarten.

```{r}
summary(STARdata$readk)
```

The students were blocked into schools via the `schoolidk` variable:

```{r}
length(unique(STARdata$schoolidk))
head(table(STARdata$schoolidk))
```

Students were the assigned units, so we need a unique identifier per student. If
this does not currently exist, it can easily be generated:

```{r}
STARdata$studentid <- seq_len(nrow(STARdata))
head(STARdata$studentid)
```

## A Basic Example

### Defining the `Design`

The three `_design` functions (`rct_design()`, `obj_design()`, and
`rd_design()`) operate similarly. The first argument is the most important, and
encodes all the design information through the use of a formula. The left-hand
side of the formula identifies the treatment variable. The right-hand side of
the formula consists of the following potential pieces of information:

1. `unit_of_assignment()`: This identifies the variable(s) which indicate the
   units of assignment. This is required for all designs. The alias `uoa()` can
   be used in its place.
2. `block()`: The identifies the variable(s) which contain block information.
   Optional.
3. `forcing()`: In regression discontinuity designs (`rd_design()`), this
   identifies the variable(s) which contain forcing information.

To define a `Design` in our example:

```{r}
des <- obs_design(starkbinary ~ unit_of_assignment(studentid) + block(schoolidk),
                  data = STARdata, na.fail = FALSE)
summary(des)
```

Should more than one variable be needed to identify the unit of assignment,
block, or forcing, they can be included. For example, perhaps `schoolidk` may be
unique within district, but potentially not unique across districts. Then we'd
use something like `block(districtid, schoolidk)` in the `_design` function.

### Estimating the treatment effect

The main function for estimating treatment effects is the `lmitt()` function. It
takes in three main required arguments:

1. A formula specifying the outcome and the desired treatment effect.
2. The data set containing the outcome information.
3. A [`Design`](#defining-the-design).

Note that the data set does **not** need to be the same data set which generated
the `Design`; it does however need to include the same variables to identify the
units of assignment. (If the variable names differ, the `by=` argument can be
used to link them, though we recommend renaming to reduce the likelihood of
issues.)

For example, you may have one dataset containing school-level information, and a
separate dataset containing student-level information. Assume school is the unit
of assignment. While you could of course merge those two data-sets,
**propertee** can instead use the school-level data to define the `Design`, and
the student-level data to estimate the treatment effect.

The formula entering `lmitt()` can take on one of two forms:

```r
y ~ 1
```

will estimate the main treatment effect on outcome variable `y`, and

```r
y ~ x
```

will estimate subgroup specific treatment effects for each level of `x` for the
outcome `y`, if `x` is categorical. For continuous `x`, a main effect and a
treatment-`x` interaction effect is estimated.

Therefore, to estimate the treatment effect in our example, we can run:

```{r}
te <- lmitt(readk ~ 1, data = STARdata, design = des)
summary(te)
```

The data includes ethnicity; we can estimate subgroup effects by ethnicity:

```{r}
te_s <- lmitt(readk ~ ethnicity, data = STARdata, design = des)
summary(te_s)
```

### Including design weights

Design weights can be easily included in this estimation. **propertee** supports
average treatment effect (ATE) and effect of the treatment on the treated (ETT)
weights.

To include one of the weights, simply include the `weights = "ate"` or
`weights = "ett"` argument to `lmitt()`:

```{r}
lmitt(readk ~ 1, data = STARdata, design = des, weights = "ate")
lmitt(readk ~ 1, data = STARdata, design = des, weights = "ett")
```

Internally, these call the `ate()` or `ett()` functions which can be used
directly.

```{r}
head(ate(des, data = STARdata))
```

When included inside `lmitt()`, you do not need to specify any additional
arguments to `ate()` or `ett()`, enabling easy functions of weights. For example
if you had some other weight variable, say `wgt`, you could include `weights =
wgt*ate()` in the `lmitt()` call.

### Covariance Adjustment models

By itself, `lmitt()` does not allow for other covariates; e.g. something like
`lmitt(y ~ 1 + control_var,...` will fail. To adjust for covariates, a separate
covariate model should be fit. Any model which supports a `predict()` function
should work.

```{r}
camod <- lm(readk ~ gender + birth + lunchk, data = STARdata)
```

The `cov_adj()` function can be used to process the covariance adjustment model
and produce the required values; and its output can be passed as an `offset=`.

```{r}
lmitt(readk ~ 1, data = STARdata, design = des,
      weights = "ate", offset = cov_adj(camod))
```

Similarly to the weight functions, `cov_adj()` attempts to locate the correct
arguments (in this case, mainly the `data=` argument) to use in the model
command; while `cov_adj()` does fall back to using the data which is in the
covariance model, its safer to use the `newdata=` argument if calling
`cov_adj()` outside of the model.

```{r}
head(cov_adj(camod, newdata = STARdata))
```

Also, similarly to weights, `cov_adj()` can be used in normal modeling commands
as well.

```{r}
lm(readk ~ starkbinary, data = STARdata, weights = ate(des),
   offset = cov_adj(camod))
```

### Absorbing Blocks

If fixed effects for blocks are desired, which can be absorbed away to avoid
estimating, the `absorb=TRUE` argument can be passed.

```{r}
lmitt(readk ~ 1, data = STARdata, design = des, absorb = TRUE)
```
