---
title: "Introduction to the 'flexida' package"
author: "flexida authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the 'flexida' package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(flexida)
```

# Main Features

The `{flexida}` package offers a two step approach to treatment effect
estimation.

1. Generate a `Design` object which encodes the study design, including the unit
   of assignment, treatment status, and optionally block information.
2. Estimate a treatment effect, accounting for the design information.

The treatment effect estimation can optionally include proper weighting (for
either average treatment effect or effect of the treatment on the treated), or
can include a externally fit covariance adjustment model.

While both the weights and and covariance adjustment can be used in a stock R
model (e.g. `lm()`), `{flexida}` offers the `lmitt()` function which can be used
to directly estimate the desired treatment effect, along with properly
calculated standard errors accounting for the design features and covariance
adjustment first stage model.

# Example Data

The example dataset comes from the state of Tennessee's Student-Teacher
Achievement Ratio (STAR) experiment. Students were randomly assigned to three
possible classroom conditions: small (13 to 17 students per teacher), regular
class (22 to 25 students per teacher), and regular-with-aide class (22 to 25
students with a full-time teacher's aide).

```{r}
data(STARdata)
table(STARdata$stark)
```

For simplicity for this first example, we will examine a single binary
treatment - "small" classrooms versus "regular" and "regular+aide" classrooms.

```{r}
STARdata$starkbinary <- STARdata$stark == "small"
table(STARdata$starkbinary)
```
After this basic example, we will see how `{flexida}` makes it easy to handle
non-binary treatment variables by introducing `dichotomy`s.

The outcome of interest is a reading score at the end of kindergarten.

```{r}
summary(STARdata$readk)
```

The students were blocked into schools via the `schoolidk` variable:

```{r}
length(unique(STARdata$schoolidk))
head(table(STARdata$schoolidk))
```

Students were the assigned units, so we need a unique identifier per student. If
this does not currently exist, it can easily be generated:

```{r}
STARdata$studentid <- seq_len(nrow(STARdata))
head(STARdata$studentid)
```

# A Basic Example

## Defining the Design

The three `_design` functions (`rct_design()`, `obj_design()`, and
`rd_design()`) operate very similarly. The first argument is the most important,
and encodes all the design information through the use of a formula. The
left-hand side of the formula identifies the treatment variable. The right-hand
side of the formula consists of the following potential pieces of information:

1. `unit_of_assignment()`: This identifies the variable(s) which indicate the
   units of assignment. This is required for all designs. The alias `uoa()` can
   be used in its place.
2. `block()`: The identifies the variable(s) which contain block information.
   Optional.
3. `forcing()`: In regression discontinuity designs (`rd_design()`), this
   identifies the variable(s) which contain forcing information.

Let's see how to define the design object.

```{r}
des <- obs_design(starkbinary ~ unit_of_assignment(studentid) + block(schoolidk),
                  data = STARdata)
summary(des)
```

Should more than one variable be needed to identify the unit of assignment,
block, or forcing, they can be included. For example, perhaps `schoolidk` may be
unique within district, but potentially not unique across districts. Then we'd
use something like `block(districtid, schoolidk)` in the `_design` function.

## Estimating the treatment effect

The main function for estimating treatment effects is the `lmitt()` function. It
takes in three main required arguments:

1. A formula specifying the outcome and the desired treatment effect.
2. The data set containing the outcome information.
3. A `Design`.

Note that the data set does **not** need to be the same data set which generated
the `Design`; it does however need to include the same variables to identify the
units of assignment. (Note that if the variable names differ, the `by=` argument
can be used to link them.) For example, treatment may be assigned at a school
level, so the data set creating the `Design` is school-level. However, the
outcome could be at the student level. As long as the student-level data had
school identifiers, `{flexida}` can match them up.

The formula entering `lmitt()` can take on one of two forms:

```r
y ~ 1
```

will estimate the main treatment effect on outcome variable `y`, and

```r
y ~ sbgrp
```

will estimate subgroup specific treatment effects for each level of `sbgrp` for
the outcome `y`.

(We hope in the future to support `y ~ 1 + sbgrp` to simultaneously estimate
both main and subgroup effects; however for now you can fit each model
separately.)

Therefore, to estimate the treatment effect in our example, we can run:

```{r}
lmitt(readk ~ 1, data = STARdata, design = des)
```

The data includes ethnicity; we can estimate subgroup effects by ethnicity.

```{r}
lmitt(readk ~ ethnicity, data = STARdata, design = des)
```

## Including design weights

Design weights can be easily included in this estimation. `{flexida}` supports
average treatment effect (ATE) and effect of the treatment on the treated (ETT)
weights.

To include one of the weights, simply include the `weights = "ate"` or
`weights = "ett"` argument to `lmitt()`:

```{r}
lmitt(readk ~ 1, data = STARdata, design = des, weights = "ate")
lmitt(readk ~ 1, data = STARdata, design = des, weights = "ett")
```

Internally, these call the `ate()` or `ett()` functions which can be used
directly.

```{r}
lmitt(readk ~ 1, data = STARdata, design = des, weights = ate())
```

This enables easy combination of weights; for example if you had some other
weight contained with-in a variable, say `wgt`, you could include `weights =
wgt*ate()` in the `lmitt()` call.

The weight functions also function within normal model calls such as `lm()` or
`glm()`.

```{r}
lm(readk ~ starkbinary, data = STARdata, weights = ate(des))
```

Note the neccessity to include the `Design` as an argument to `ate()`; `ate()`
and `ett()` attempt to locate both the `Design` and `data` inside their call,
however if needed, both can be manually included. For example, to obtain the
weights indepndently on the model:

```{r}
head(ate(des, data = STARdata))
```

## Covariance Adjustment models

By itself, `lmitt()` does not allow for other covariates; e.g. something like
`lmitt(y ~ 1 + control_var,...` will fail. To adjust for covariates, a separate
covariate model should be fit. Any model which supports a `predict()` function
will work.

```{r}
camod <- lm(readk ~ gender + birth + lunchk, data = STARdata)
```

The `cov_adj()` function can be used to process the covariance adjustment model
and produce the required values; and its output can be passed as an `offset=`.

```{r}
lmitt(readk ~ 1, data = STARdata, design = des,
      weights = "ate", offset = cov_adj(camod))
```

Similarly to the weight functions, `cov_adj()` attempts to locate the correct
dataset to use in the model command; while `cov_adj()` does fall back to using
the data which is in the covariance model, its safer to use the `newdata=`
argument if calling `cov_adj()` outside of the model.

```{r}
head(cov_adj(camod, newdata = STARdata))
```

Also, similarly to weights, `cov_adj()` can be used in normal modeling commands
as well.

```{r}
lm(readk ~ starkbinary, data = STARdata, weights = ate(des),
   offset = cov_adj(camod))
```

## Absorbing Blocks

If it is desired to include fixed effects for blocks, which can be absorbed away
to avoid estimating, the `absorb=TRUE` argument can be passed.

```{r}
lmitt(readk ~ 1, data = STARdata, design = des, absorb = TRUE)
```

# Standard error and confidence bound calculation


# Additional topics

## Non-binary treatment variables

In the example above, we manually created a binary treatment variable,
`starkbinary` out of a three-level categorical variable. `Design`s can instead
handle non-binary treatment variables. (Note that factor variables, while able
to be included in a `*_design` call, have their labels converted to numeric or
character as appropriate. We therefore recommend doing this conversion a priori
so you can control this conversion and ensure it is appropriate.)

```{r}
STARdata$stark <- as.character(STARdata$stark)
des <- obs_design(stark ~ unit_of_assignment(studentid) + block(schoolidk),
                  data = STARdata)
summary(des)
```

Note however that some features (notably weight generation and proper covariance
estimates) are not fully supported in cases with non-binary treatment. While we
could manually create several binary variables based upon the non-binary
treatment variables, we can instead use the `dichotomy` feature to inform the
`Design` about the current binary comparison we are examing.

A `dichotomy` is given as a formula object, where the left hand side evaluates
to a logical statement about the treatment variable identifying the "treated"
group, and the right hand side evaluates to a logical statement identifying the
"control" group. For example, if we wanted to examine the treatment effect of
"small" versus "regular", the following `dichotomy` would provide this
information:

```r
stark == "small" ~ stark == "regular"
```

The "small" versus pooled "regular"/"regular+aide" can be expressed in any
several ways:

```r
stark == "small" ~ stark == "regular" | stark == "regular+aide"
stark == "small" ~ stark != "small"
stark == "small" ~ .
. ~ stark == "regular" | stark == "regular+aide"

```

The "`.`" can be used in exactly one side, and implies that that group is
defined as all units not assigned to the other group.

You can attach a `dichotomy` to a `Design` with the `dichotomy()` function:

```{r}
des2 <- des
dichotomy(des2) <- stark == "small" ~ .
des2
```

Once a `dichotomy` is attached to a `Design`, its equivalent for all intents and
purposes to if you had used a binary treatment variable when creating the
`Design`. (The `*_design` functions take a `dichotomy=` argument so you can
create a `Design` and attach a `dichotomy in one step.)

Alternatively, the `ate()` and `ett()` functions, for generating average
treatment effect weights and effect of the treatment on the treated weights
respectively, take an optional `dichotomy=`
