---
title: "Regression Discontinuity Designs"
author: "Adam Sales"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regression Discontinuity Designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
```

```{r}
library(flexida)
```

## Background

### Regression Discontinuity Designs

### Analyzing RDDs with ANCOVA

### The `flexida` Approach to RDDs

### RDD Specification Tests


## Data and Design

The data were compiled by Ludwig & Miller (2007; henceforth LM), largely from the National Archives and Records Administration and Vital Statistics, and were cleaned and provided by the `RDHonest` package for R.

At the onset of the Head Start preschool program in 1965, the Office of Economic Opportunity provided technical support in applying for funding to the 300 US counties with the highest 1960 poverty rates. LM show that this assistance paid off, in that couties with poverty rates just above the cutoff received substantially more Head Start funding than those with poverty rates just below the cutoff. LM then sought to estimate the impact of this additional funding on two county-level outcomes that may plausibly affected by Head Start: child mortality due to health risks addressed by the Head Start program, and high school completion rates. 

```{r, eval=FALSE}
devtools::install_github('kolesarm/RDHonest')
```

```{r, results = 'hide', message = FALSE, error = FALSE, warning = FALSE}
library(RDHonest)
data(headst)
```

