---
title: "Applying Dichotomies to Non-Binary Treatments"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include = FALSE}
devtools::load_all()
```

## Binary Treatment

When creating a Design, handling binary treatment variables is straightforward:

```{r}
des1 <- rct_design(z ~ cluster(cid1, cid2), data = simdata)
head(ate(des1, data = simdata))
```

## Non-binary Treatment

However, while Design objects support non-binary treatment variables, attempting
to utilize those non-binary treatment variables will result in an error:

```{r, error= TRUE}
head(simdata$dose)
des2 <- rct_design(dose ~ cluster(cid1, cid2), data = simdata)
ate(des2, data = simdata)
```

### Applying a `dichotomy`

Instead, you need to define a `dichotomy`. A `dichotomy` is presented as a
formula, where the left-hand side is a logical statement defining inclusion in
the treatment group, and the rihgt-hand side is a logical statement defining
inclusion in the control group. For example,

```{r, eval = FALSE}
dose > 200 ~ dose <= 200
```

defines that all units of assignment with `dose` above 200 are treated units,
and all units of assignment with `dose` of 200 or below are control units.

A `.` can be used to define either group as the inverse of the other. For
example, the above dichotomy could be defined as either of

```{r, eval = FALSE}
dose > 200 ~ .
. ~ dose <= 200
```

Note that all units need to be assigned to treatment or control; any units who
are not captured in either logical statement are assumed to be excluded from the
trial.

```{r, eval = FALSE}
dose > 200 ~ dose <= 200 & dose >= 100
```

In this `dichotomy`, only units of assignment above 100 are considered in the
trial.

Let's see these passed to the `ate()`/`ett()` functions:

```{r}
head(ate(des2, data = simdata,
         dichotomy = dose > 200 ~ dose <= 200))
head(ate(des2, data = simdata,
         dichotomy = dose > 200 ~ .))
head(ate(des2, data = simdata,
         dichotomy = .~ dose <= 200))
head(ate(des2, data = simdata,
         dose > 200 ~ dose <= 200 & dose >= 100))
```

The first four units of assignment have `dose` of 50, notice their corresponding
weight of 0 in the last example.

Alternatively, we could place the dichotomy directly in the `Design` object
instead.

```{r}
dichotomy(des2) <- dose > 200 ~ dose <= 200 & dose >= 100
head(ate(des2, data = simdata))
des3 <- rct_design(dose ~ cluster(cid1, cid2), data = simdata,
                   dichotomy = dose > 200 ~ dose <= 200 & dose >= 100)
head(ate(des3, data = simdata))
```
### Storage of non-binary treatment

When a design's treatment is dichotomized, we keep the orignal treatment in
place, but add a slot for the `dichotomy`:

```{r}
des3
```

We can obtain the `dichotomy` with the `dichotomy()` function, or apply it to
the treatment with the `treatment()` function and it's `binary=` argument.

```{r}
dichotomy(des3)
treatment(des3)
treatment(des3, binary = TRUE)
```

### Helper Functions

There are three helper functions:

- `has_binary_treatment()` identifies whether a `Design` has a binary treatment
  variable.
- `is_dichotomized()` identifies whether a `Design` has a `dichotomy`
  applied to it.
- `is_binary_or_dichotomized()` identifies whether either a `Design` is binary
  **or** has a `dichotomy` applied to it.

```{r}
c(is_dichotomized(des1),
  has_binary_treatment(des1),
  is_binary_or_dichotomized(des1))
c(is_dichotomized(des2),
  has_binary_treatment(des2),
  is_binary_or_dichotomized(des2))
```
